Distrivia: A Distributed Trivia Game

### 2011-04-22 papers/phase-two/paper.tex
Minor tweaks. Caps Android, fix bcrypt spellling add Monit

\section {What We've Done}
Much of our system has been implemented and is in the debugging phase. The server apps, web client, and android client are nearly complete with some revisions happening as necessary. We have also started development on an application for the iPhone platform. The extent of what has been finished for each part of our system is explained below.

@@

\subsection {Servers}
Currently, we have 6 servers running on Amazon's Elastic Compute Cloud \cite{aec}. One server is being used as a web-monitor
host. We are able to login to the web-monitor to see the status of all our
servers, databases, and load sharers. This will help us manage if the services
are running and allow us to track down any problems, if they arise. Three
servers are all running our server application and hosting the databases.
We also have two final servers that we have setup as hot spares in the event
of partial or total failure.
Each server contains the server application that responds to clients as well as
the Riak \cite{riak} databases. Riak auto-synchronizes the databases between the 5 servers.
Note that the hot spares are setup to be part of the database cluster so they
have a full copy of the database on hand, however they don't accept any reads
or writes directly. This allows for them to immediately be ready for read/writes in the event that the original servers are down.
With this configuration if we were to add a server, we would just need to update 
Riak and it would sync the databases to any additional servers as well. These servers,
therefore, all contain the same information. If any servers go down, the distributed
system would still run fine. We could have all servers go down except one and still
be reliably handling requests.

@@

\subsection {Clients}
Our focus for clients has been on a web-based client and an android application.
Both clients have most of the the workings completed. The web-based client will
be run from desktop platforms and laptop platforms. The android client will run
on any android phone running OS 2.1 or higher and does not use any unique services that require specific
hardware. Both clients are nearly fully-functional. They can, currently, login/register
players, view leaderboards, join public games, and compete in public games. We are actively adding features to both clients, daily.
We have also started development on an iPhone application for distrivia. This client does not yet communicate with the server, but 
it does implement nearly all the views seen in the other clients. We have been setting up the network communication for the iPhone
platform as well.

\subsection {Messages}
We are currently using the https protocol for our messages. This provides a secure
and usable messaging system that we have designed our client and server APIs to use.
The actual messages are sent and received in two way's. Initially a user session key is
generated using the UUID \cite{uuid} algorithm and returned to the user on successful
authentication. This session key is then passed back to the server with every in coming
message. Messages are sent to the server using POST and GET request methods supported
by the HTTP protocol. Responses from the server are in the form of JSON \cite{json} objects.
JSON is a small, easily parseable format, representing objects in a human readable format as
they would be represented in JavaScript. With the web client the JSON object returned from
the server is instantly usable, however for the android client we have to parse the JSON object
into a native Java object. The Android SDK provides libraries to do this, located in the
org.json \cite{orgjson} package name space.

\section {What's Left}
\subsection {Server}
The server is nearly completed. The server application just needs to have some tweaks and further testing.
\subsection {Web Client}
A lot of the main features of the web client are completed.
However, there are still a few remaining features that need to be implemented.
The only major feature remaining is private games, which are largely unstarted, although it will reuse most of the public game code.
In addition to this, buttons should be disabled when clicked to prevent users from spamming the server with repeat messages.
\subsection {Android}
Private gameplay is the largest thing left for the android client. Besides this we've just been fixing bugs and cleaning up the UI.
\subsection {iPhone}
The iPhone application just started development in the last couple days. We have built most of the UI but need to connect all of the actions
to the server. We will need to implement everything, including login/registration, joining public/private games, viewing leaderboards,
and competing in rounds.

@@

Our team is taking a very proactive stance towards security. All of our machines are properly firewalled so that
outside clients can not authenticate to our database cluster. Also we have implemented SSL through our entire infrastructure.
Our load sharer's automatically redirect http:// to https:// and once you are in https:// at the load sharer, SSL continues all the
way back to our actual web application back end serving responses over a SSL wrapped socket. This way no passwords are ever
sent in clear text. Our group even found a SSL certificate authority who was willing to sign our SSL certificate for free. This gives
us actual fully functional, authenticatable SSL infrastructure. In order to make sure that our system properly secures user information,
we stuck with the industry standard of storing a hashed version of the users password instead of the password itself. We decided on using
the cryptographically secure hashing function designed for passwords known as bcyrpt \cite{bcrypt}.

>>>>

\section {What We've Done}
Much of our system has been implemented and is in the debugging phase.
The server application, web client, and Android client are nearly complete with some revisions happening as necessary.
We have also started development on an application for the iPhone platform.
The extent of what has been finished for each part of our system is explained below.

@@

\subsection {Servers}
Currently, we have 6 servers running on Amazon's Elastic Compute Cloud \cite{aec}. One server is being used as a web-monitor
host running Monit\cite{monit}. We are able to login to the web-monitor to see the status of all our
servers, databases, and load sharers. This will help us manage whether or not the services
are running and allow us to track down any problems, if they arise. Three
servers are all running our server application and hosting the databases.
We also have two final servers that we have setup as hot spares in the event
of partial or total failure.
Each server contains the server application that responds to clients as well as
the Riak \cite{riak} databases. Riak auto-synchronizes the databases between the 5 servers.
Note that the hot spares are setup to be part of the database cluster so they
have a full copy of the database on hand, however they don't accept any reads
or writes directly. This allows for them to immediately be ready for read/writes in the event that the original servers are down.
With this configuration if we were to add a server, we would just need to update 
Riak and it would sync the databases to any additional servers as well. These servers,
therefore, all contain the same information. If any servers go down, the distributed
system would still run fine. We could have all servers go down except one and still
be reliably handling requests.

@@

\subsection {Clients}
Our focus for clients has been on a web-based client and an Android application.
Both clients have most of the workings completed. The web-based client will
be run from desktop platforms and laptop platforms. The Android client will run
on any Android phone running OS 2.1 or higher and does not use any unique services that require specific
hardware. Both clients are nearly fully-functional. They can, currently, login/register
players, view leaderboards, join public games, and compete in public games. We are actively adding features to both clients, daily.
We have also started development on an iPhone application for distrivia. This client does not yet communicate with the server, but 
it does implement nearly all the views seen in the other clients. We have been setting up the network communication for the iPhone
platform as well.

\subsection {Messages}
We are currently using the https protocol for our messages. This provides a secure
and usable messaging system that we have designed our client and server APIs to use.
The actual messages are sent and received in two way's. Initially a user session key is
generated using the UUID \cite{uuid} algorithm and returned to the user on successful
authentication. This session key is then passed back to the server with every in coming
message. Messages are sent to the server using POST and GET request methods supported
by the HTTP protocol. Responses from the server are in the form of JSON \cite{json} objects.
JSON is a small, easily parseable format, representing objects in a human readable format as
they would be represented in JavaScript. With the web client the JSON object returned from
the server is instantly usable, however for the Android client we have to parse the JSON object
into a native Java object. The Android SDK provides libraries to do this, located in the
org.json \cite{orgjson} package name space.

\section {What's Left}
\subsection {Server}
The server is nearly completed. The server application just needs to have some tweaks and further testing.
\subsection {Web Client}
A lot of the main features of the web client are completed.
However, there are still a few remaining features that need to be implemented.
The only major feature remaining is private games, which are largely unstarted, although it will reuse most of the public game code.
In addition to this, buttons should be disabled when clicked to prevent users from spamming the server with repeat messages.
\subsection {Android}
Private gameplay is the largest thing left for the Android client. Besides this we've just been fixing bugs and cleaning up the UI.
\subsection {iPhone}
The iPhone application just started development in the last couple days. We have built most of the UI but need to connect all of the actions
to the server. We will need to implement everything, including login/registration, joining public/private games, viewing leaderboards,
and competing in rounds.

@@

Our team is taking a very proactive stance towards security. All of our machines are properly firewalled so that
outside clients can not authenticate to our database cluster. Also we have implemented SSL through our entire infrastructure.
Our load sharer's automatically redirect http:// to https:// and once you are in https:// at the load sharer, SSL continues all the
way back to our actual web application back end serving responses over a SSL wrapped socket. This way no passwords are ever
sent in clear text. Our group even found a SSL certificate authority who was willing to sign our SSL certificate for free. This gives
us actual fully functional, authenticatable SSL infrastructure. In order to make sure that our system properly secures user information,
we stuck with the industry standard of storing a hashed version of the users password instead of the password itself. We decided on using
the cryptographically secure hashing function designed for passwords known as bcrypt \cite{bcrypt}.
